Determinism, Complexity, and Predictability in Computer Performance

### 2014-03-06 intro.tex
add comment about redundancy to intro

Unfortunately, as a corollary of the undecidability of the halting problem: there cannot exist \emph{one} universally ideal forecasting scheme for even completely noise-free deterministic time series\cite{weigend-book}, let alone an arbitrary time series. This naturally leads to an important and hard question: given a complicated real-valued time series, does their exist a forecast model for that time series that can utilize the information being transmitted from past to future (if such information exists)? A first step in answering this question is to answer: is it possible to reliably quantify where on the spectrum of complexity the time series lies? i.e, when little or nothing is known about the underlying system's generating process (e.g., linear, nonlinear, deterministic, stochastic, stationary, non-stationary etc.) can it be inferred without a model if the signal is too \emph{complex} to effectively forecast using a prediction strategy based on predictive structure (temporal forward information transfer)? We rigorously define this \emph{complexity spectrum} as well as what is meant by \emph{low, medium and high} in Section \ref{sec:meaComplex}.

>>>>

Unfortunately, as a corollary of the undecidability of the halting problem: there cannot exist \emph{one} universally ideal forecasting scheme for even completely noise-free deterministic time series\cite{weigend-book}, let alone an arbitrary time series. This naturally leads to an important and hard question: given a complicated real-valued time series, does their exist a forecast model for that time series that can utilize the information being transmitted from past to future (if such information exists)? A first step in answering this question is to answer: is it possible to reliably quantify where on the spectrum of complexity the time series lies? i.e, when little or nothing is known about the underlying system's generating process (e.g., linear, nonlinear, deterministic, stochastic, stationary, non-stationary etc.) can it be inferred without a model if the signal is too \emph{complex} to effectively forecast? Since the information in an observation can be partitioned into two pieces --- redundancy and entropy generation\cite{crutchfield2003} --- our goal is to analyze how the amount of predictive structure present in a signal, as measured by the redundancy, relates to the performance of prediction strategies, as well as what is meant by \emph{low, medium and high} complexity in Section \ref{sec:meaComplex}.
